---
output:
  html_document:
    fig_height: 5
    fig_width: 5
    fig_position: center
  always_allow_html: true
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                message = FALSE, warning = FALSE, 
                fig.align = "center")
options(htmltools.dir.version = FALSE)

```

---
title: " Bioerosion Monitoring Unit QAQC"
author: "Nicole Besemer"
date: "`r format(Sys.time(), '%d %B, %Y')`"

---

```{r setup2, include=FALSE}
library(knitr)
library(lubridate)
library(dplyr)
library(tools)
library(scales)
library(ggplot2)
library(kableExtra)
library(ggpubr)
library(webshot)
library(readr)
library(timechange)

knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, echo = FALSE, cache = TRUE, autodep = TRUE, cache.comments = FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE, cache.lazy = FALSE, results = "asis")
```

Atlantic 2023 Data

```{r load data, echo=FALSE, include=FALSE}


data<-read_csv("~/Documents/R/bmuQAQC/NCRMP_BMU_Atlantic2022.csv")
head(data)


```


```{r set_dates, echo=FALSE}

data$Deployment = mdy(data$Deployment)
data$Recovery = mdy(data$Recovery)
data$recoveryYear<- as.numeric(format(data$Recovery, "%Y"))


```

```{r set bounds}

# Initiate the data frame and name the rows and columns
vars =  c(

  "Latitude",
  "Longitude",
  
  "PreWeightBlock",
  "PreWeightEpoxied",
  "PreVolume",
  "PreRWDensity",
  "PreCTDensity",
  
  "PostWeightDirty",
  "PostWeightClean",
  "PostVolume",
  "PostCTDensity",
  
  "PostCTMacroboring",
  "PostCTMacroboringRate",
  "PostCTGrazing",
  "PostCTGrazingRate",
  "PostCTAccretion",
  "PostCTAccretionRate",
  "recoveryYear"
)

bounds = as.data.frame(matrix(nrow = 2, ncol = length(vars)))
colnames(bounds) = vars
rownames(bounds) = c("Min", "Max")



bounds$Latitude = c(min(data$Latitude, na.rm = TRUE), max(data$Latitude, na.rm = TRUE))
bounds$Longitude = c(min(data$Longitude, na.rm = TRUE), max(data$Longitude, na.rm = TRUE))


bounds$PreWeightBlock = c(min(data$PreWeightBlock, na.rm = TRUE), max(data$PreWeightBlock, na.rm = TRUE))
bounds$PreWeightEpoxied = c(min(data$PreWeightEpoxied, na.rm = TRUE), max(data$PreWeightEpoxied, na.rm = TRUE))
bounds$PreVolume = c(min(data$PreVolume, na.rm = TRUE), max(data$PreVolume, na.rm = TRUE))
bounds$PreRWDensity = c(min(data$PreRWDensity, na.rm = TRUE), max(data$PreRWDensity, na.rm = TRUE))
bounds$PreCTDensity = c(min(data$PreCTDensity, na.rm = TRUE), max(data$PreCTDensity, na.rm = TRUE))


bounds$PostWeightDirty = c(min(data$PostWeightDirty, na.rm = TRUE), max(data$PostWeightDirty, na.rm = TRUE))
bounds$PostWeightClean = c(min(data$PostWeightClean, na.rm = TRUE), max(data$PostWeightClean, na.rm = TRUE))
bounds$PostVolume = c(min(data$PostVolume, na.rm = TRUE), max(data$PostVolume, na.rm = TRUE))
bounds$PostCTDensity = c(min(data$PostCTDensity, na.rm = TRUE), max(data$PostCTDensity, na.rm = TRUE))


bounds$PostCTMacroboring = c(min(data$PostCTMacroboring, na.rm = TRUE), max(data$PostCTMacroboring, na.rm = TRUE))
bounds$PostCTMacroboringRate = c(min(data$PostCTMacroboringRate, na.rm = TRUE), max(data$PostCTMacroboringRate, na.rm = TRUE))
bounds$PostCTGrazing = c(min(data$PostCTGrazing, na.rm = TRUE), max(data$PostCTGrazing, na.rm = TRUE))
bounds$PostCTGrazingRate = c(min(data$PostCTGrazingRate, na.rm = TRUE), max(data$PostCTGrazingRate, na.rm = TRUE))
bounds$PostCTAccretion = c(min(data$PostCTAccretion, na.rm = TRUE), max(data$PostCTAccretion, na.rm = TRUE))
bounds$PostCTAccretionRate = c(min(data$PostCTAccretionRate, na.rm = TRUE), max(data$PostCTAccretionRate, na.rm = TRUE))
bounds$recoveryYear = c(min(data$recoveryYear, na.rm = TRUE), max(data$recoveryYear, na.rm = TRUE))

bounds<-round(bounds, digits=3)
bounds$recoveryYear<-as.integer(bounds$recoveryYear)


bounds = t(bounds)

```


```{r summary}
# Number of samples per island
summary.samples = data %>%
  group_by(RegionCode, recoveryYear) %>%
  summarize(
    totalSamples = length(tagNum),
    deplomentDuration = mean(Duration, na.rm = TRUE),
    numberRecovered = length(Fate == 'SUCCESSFULY RECOVERED'))
```



QC run at: `r now(tzone = "EST")` EST

Number of samples in data file: `r nrow(data)`




## Data extents and ranges

Macroboring, macroboring rate, grazing and grazing rate represent negative values

```{r bounds}
kable(bounds, format = "html", booktabs = TRUE, longtable = TRUE, omit.na = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position", "repeat_header"), full_width = FALSE)
```


## Number of samples in report


```{r}
kable(summary.samples, format = "html", booktabs = TRUE, longtable = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position", "repeat_header"), full_width = FALSE)
```



## QAQC Plots


### Change in Volume vs Change in Weight

Should have a linear relationship


```{r volumeVSmass, echo=FALSE, fig.width = 6, fig.height = 4}

plot1<-ggplot(data = data) +
  geom_point(aes(x = (PostVolume-PreVolume), 
                y = (PostWeightClean-PreWeightEpoxied)),
             size=3, colour = "black")+
  scale_x_continuous(name=bquote('volDifference'~(cm^3)))+
  scale_y_continuous(name = "weightDifference (g)")

plot1


```





###  Macroboring by bmu

Check for non-negatives and any extreme outliers




```{r plot macroboring, echo=FALSE, fig.width = 6, fig.height = 4}
plot2<-ggplot(data = data) +
  geom_point(aes(x = tagNum, 
                y = PostCTMacroboring*-1),
             size=3, colour = "red")+
  scale_y_continuous(name = bquote('PostCTMacroboring'~(cm^3)))
plot2
```



### Grazing by BMU

There should be no positive values


```{r plotgrazing, echo=FALSE, fig.width = 6, fig.height = 4}
plot3<-ggplot(data = data) +
  geom_point(aes(x = tagNum, 
                y = PostCTGrazing*-1),
             size=3, colour = 'red')+
  scale_y_continuous(name = bquote('PostCTGrazing'~(cm^3)))
plot3


```



### Accretion by BMU
```{r accretion, echo=FALSE, fig.width = 6, fig.height = 4}
plot4<-ggplot(data = data) +
  geom_point(aes(x = tagNum, 
                y = PostCTAccretion),
             size=3, colour = "darkgreen")+
  scale_y_continuous(name = bquote('PostCTAccretion'~(cm^3)))
plot4
```








### Grazing vs Accretion



```{r grazingVSaccretion, echo=FALSE, fig.width = 6, fig.height = 4}
plot5<-ggplot(data = data) +
  geom_point(aes(x = PostCTAccretion, 
                y = (PostCTGrazing*-1)),
             size=3, colour = "black")+
  scale_x_continuous(name=bquote('PostCTAccretion'~(cm^3)))+
  scale_y_continuous(name = bquote('PostCTGrazing'~(cm^3)))
plot5
```




### Net Accretion vs Net Erosion



```{r netAccretion vs net erosion, echo=FALSE, fig.width = 6, fig.height = 4}

plot6<-ggplot(data = data) +
  geom_point(aes(x = (PostCTMacroboring+PostCTGrazing), 
                y = (PostCTAccretion)),
             size=3, colour = "black")+
  scale_x_continuous(name = bquote('netErosion'~(cm^3)))+
  scale_y_continuous(name = bquote('netAccretion'~(cm^3)))

plot6


```


### Bioerosion Rate

combination of macroboring and grazing (lack of pre scan information for STT and STJ means we cannot calculate the erosion rate for these locations)


```{r bioerosion rate, echo=FALSE, fig.width = 5, fig.height = 5}

plot7<- ggplot(data=data) + 
  geom_boxplot(aes(x=RegionCode, y=((PostCTGrazingRate+PostCTMacroboringRate)*-1)))+
  labs(y = bquote('Erosion Rate'~(cm^3/year)))+ geom_point(aes(x=RegionCode, y=((PostCTGrazingRate+PostCTMacroboringRate)*-1)), shape=16, size=3)

plot7

```

### Accretion Rate

Florida keys samples have a noticeably higher accretion rate. There is a higher sample size and some were left out significantly longer than others. Should like into how deployment time affects accretion rates. 


```{r accretionRate, echo=FALSE, , fig.width = 5, fig.height = 5}

plot8<- ggplot(data=data) + 
  geom_boxplot(aes(x=RegionCode, y=(PostCTAccretionRate)))+
  labs(x="Region", y = bquote('Accretion Rate'~(cm^3/year)))+ geom_point(aes(x=RegionCode, y=(PostCTAccretionRate)), shape=16, size=3)

plot8


```


### All BMU Accretion and Erosion

```{r bardata, include=FALSE}
library(tidyr)
data2<-select(data,RegionCode, tagNum, PostCTGrazing, PostCTMacroboring, PostCTAccretion)
data2$PostCTMacroboring<-data2$PostCTMacroboring*-1
data2$PostCTGrazing<-data2$PostCTGrazing*-1
head(data2)


data3 <- gather(data2,
                   key = "bmuAttribute",
                   value = "carbAmount",
                   -tagNum, -RegionCode)
head(data3)


```


```{r bar graph, echo=FALSE, fig.width = 8, fig.height = 4}


data3$bmuAttribute <- factor(data3$bmuAttribute, levels = c("PostCTGrazing", "PostCTmacroboring", "PostCTAccretion"))
data3$carbAmount = as.numeric(as.character(data3$carbAmount))

bmubar = ggplot(data = data3, aes(x=RegionCode, y=carbAmount, fill=bmuAttribute))+
  geom_bar(stat="identity", color="black",
           position = position_dodge())+
  labs(y=expression("Amount change (cm3)"), x="BMU Region")+
  scale_fill_manual(name = "bmuAttribute", labels = c("Grazing", "Accretion", "Macroboring"), values=c("#009E73", "#56B4E0","#D55E00" ))
bmubar







```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
